Implicit yaddayadda

### 2011-10-11 papers/PLDI2012/implicitization.tex
partie disjunctive invariant

Our improvement consists in a slight modification of the SMT formula we solve in
\ref{sec:guided_multigraph}.
We introduce in this formula Boolean variables $\{d_j, 1 \leq j \leq m\}$, so
that we can easily find in the model which abstract value of the disjunction has
to be chosen to make the invariant grow.
The resulting formula that is given to the SMT solver is defined
by $g(p_i)$.
When the formula is satisfiable, we know that the index $j$ of the starting
disjunct that has to be chosen is the one for which the associate Boolean value
$d_j$ is \emph{true} in the model. Then, we can easily compute the value of 
$\sigma(j,k)$, thus know the index of the disjunct to join with.

@@

Our algorithm for computing disjunctive invariants is described in Algorithm
\ref{algo:disjunctive}. In this algorithm, instead of simply joining the two
abstract values, we use the widening operator to ensure the algorithm terminates
when using abstract domains with infinite height.

\MM{Comment cet algo se place par rapport à l'algo
  \ref{algo:combined} ? Si je comprends bien, la boucle while
  remplace l'algo \ref{algo:computepaths}, et l'initialisation au
  début remplace le début de \ref{algo:combined}. Faudrait rendre ça
  explicite.}
\begin{algorithm}[!h]
\caption{Disjunctive invariant computation with implicit paths}\label{gulwani2}
\label{algo:disjunctive}
\begin{algorithmic}[1] 
\FORALL {$k \in \{1,..,m\} \setminus \{\delta\}$}
	\STATE $X_{i,k} \gets \perp$
\ENDFOR
\STATE $X_{i,\delta} \gets I_{p_i}$
\WHILE {true}
		\STATE $res \gets SmtSolve\left[g(p_i)\right]$
	\IF {$res = unsat$}
		\STATE \textbf{break}
	\ENDIF
	\STATE Compute the path $\tau_{i,k}$ from $res$ 
	\STATE Take $j \in \{ l | d_l = true\}$ 
	\STATE $X_{i,\sigma(j,k)} \gets X_{i,\sigma(j,k)} \widening (X_{i,\sigma(j,k)} \sqcup \tau_{i,k}(X_{i,j}))$
\ENDWHILE
\end{algorithmic}
\end{algorithm}

>>>>

Our improvement consists in a modification of the SMT formula we solve in
\ref{sec:guided_multigraph}.
We introduce in this formula Boolean variables $\{d_j, 1 \leq j \leq m\}$, so
that we can easily find in the model which abstract value of the disjunction has
to be chosen to make the invariant grow.
The resulting formula that is given to the SMT solver is defined
by $g(p_i)$.
When the formula is satisfiable, we know that the index $j$ of the starting
disjunct that has to be chosen is the one for which the associate Boolean value
$d_j$ is \emph{true} in the model. Then, we can easily compute the value of 
$\sigma(j,k)$, thus know the index of the disjunct to join with.

@@

In our algorithm, the initialization of the abstract values slightly differs from
algorithm \ref{algo:combined} line~\ref{alg=X-init}, since we now have to
initialize each disjunct. Line~\ref{alg=X-init} is then replaced by:
\begin{algorithm}[!h]
\begin{algorithmic}[1] 
\FORALL {$k \in \{1,..,m\} \setminus \{\delta\}$}
	\STATE $X_{i,k} \gets \perp$
\ENDFOR
\STATE $X_{i,\delta} \gets I_{p_i}$
\end{algorithmic}
\end{algorithm}

Furthermore, the Path-focused algorithm (line~\ref{alg=pf} from algorithm
\ref{algo:combined}) is enhanced to deal with disjunctive invariants, and
is detailed in algorithm \ref{algo:disjunctive}.

The \emph{Update} function can classically assign to $X_{i,\sigma(j,k)}$ the
value $X_{i,\sigma(j,k)} \widening (X_{i,\sigma(j,k)} \sqcup
\tau_{i,k}(X_{i,j}))$, or can integrate the special treatment for self loops
proposed by \citet{Monniaux_Gonnord_SAS11}, with widening/narrowing sequence or
acceleration.

\begin{algorithm}[!h]
\caption{Disjunctive invariant computation with implicit paths}\label{gulwani2}
\label{algo:disjunctive}
\begin{algorithmic}[1] 
\WHILE {true}
		\STATE $res \gets SmtSolve\left[g(p_i)\right]$
	\IF {$res = unsat$}
		\STATE \textbf{break}
	\ENDIF
	\STATE Compute the path $\tau_{i,k}$ from $res$ 
	\STATE Take $j \in \{ l | d_l = true\}$ 
	\STATE Update($X_{i,\sigma(j,k)})$
\ENDWHILE
\end{algorithmic}
\end{algorithm}

The $\sigma$ function takes as parameters the index of the starting abstract
value, and the path we focus on. If we dynamically construct this function
during the analysis, we should store the results of the sigma function into a
compact representation, such as Algebraic Decision Diagrams.
